%\chapter{Introduction} \label{ch:intro}
\section{Motivation}
Dynamical bipedal walking has been a key objective in robotics since
its origins, due to the human curiosity about artificial
anthropomorphic beings which gave rise to the robot concept itself
\cite{Capek3R.U.R.}. As could be seen in most industrialized
countries, the industrial manipulators have found a wide adoption, and
there is little space to a major boost in the area
\cite{Yonemoto85TECHNOLOGY}. Although, the interest is shifting from
an industrial point of view to a more domestic one
\cite{Asami94Robots}, where robots can be seen as additional aids to
human daily tasks.

But, in order to accompany humans, the robot must be able to fluidly
move through all the environments in which the human can, and those
environments are devised to adapt well to anthropomorphic beings:
factories, vehicles, houses, sidewalks, and shopping malls, among
others. This way, a robot made to perform well in arbitrary
environments will have a great advantage if it is anthropomorphic, so
that it could serve well as an personal assistant
\cite{Dario01Humanoids}.


The interest on biped robotics is not only for bio-robotics
itself. Another reason to research anthropomorphic motion is the
understanding of human morphology, mechanics and control, from a
medical point of view, where robotics could serve as a testing
scenario to both theories and technologies concerning human motion
(for an example see \cite{Woo06Biomechanics}) and, probably, provide
technological aids and substitutes to body parts when an impairment is
present \cite{Hermini01Proposal}.


Another motivation to the research of biped walking is related to the
fact that anthropomorphic motion planning and control is a complex
problem that includes nonlinear and non-holonomic systems
\cite{Basdogan96Nonlinear}, complex computing tasks, adaptability to
unknown and unstructured environments \cite{Cheng00Dynamic}, among
others, and is useful to test different mechanical, electronic,
computing and control techniques applicable to diverse areas. As it is
remarked by Craig \cite{Craig89Introduction}, it should not be forgot
that the predominant dynamics algorithm for open-chain mechanisms was
developed \cite{Stepanenko6Dynamics} and refined \cite{Orin9Kinematic}
while working in biped walking problems.


Among the different problems faced in anthropomorphic motion, biped
walking is one of the most difficult because its intrinsic
instability. In contradistinction to wheeled mobile robotics and
stable legged robotics, biped robotics must allow locally unstable
motion in order to attain a fluid locomotion. Additionally, there are
several problems in biped locomotion other than stability. A bipedal
walker must be capable to choose the best path to reach an objective,
avoid obstacles, tolerate high perturbations, perform well in
unstructured environments, and move with an optimal energy
consumption.

\section{Goals}

While already a problem broadly studied, the problem of biped walking
is still open for several key goals not currently entirely satisfied,
such as optimal energy consumption, objective-based planning and
walking in less structured environments. However, there are recent,
and not so recent, notable contributions in this field obtained by
diverse methods: active control, passive dynamic walking, and
computational intelligence; which are giving a stronger basis for
major advances in the field, fueled by a strong academic and
commercial motivation.

Nonetheless, here we study a narrower problem, in which biped robotics
combines with dynamical neural neurons and evolutionary robotics to
pursue biped walking control by emergence, i.e. without the explicit
specification of those parameters required in order to achieve the
desired behavior with the control scheme chosen.

In a intuitive way, the problem studied can be broadly described as to
provide a biped robot with an optimal locomotion. Nonetheless, since
the optimality referred can be thought as a performance measure, it is
convenient to be more specific in its definition. Based on the
preliminaries in previous sections, some possible relevant goals to be
accomplished by biped locomotion are:

\begin{itemize}

\item To conserve static stability: The ability to conserve
  equilibrium while in static state or in quasi-static motion.

\item To walk conserving dynamic stability: The ability to walk, or
  locomote without leaving the contact with the floor, at velocities
  high enough to make considerable the inertial forces maintaining
  stability. That means the robot should walk indefinitely without
  falling if there are no obstacles in its way.

\item To locally minimize energy consumption: For a given path
  minimize the energy consumption of the gait pattern.

\item To react to external perturbations: Sense and compensate
  external perturbations that affect locomotion, including external
  forces and changes in environment parameters.

\item To plan gait trajectories to attain specific objectives: The
  ability to choose a suitable trajectory to reach a desired state,
  not only in terms of path, but also joint coordination.

\item To move across unstructured environments: Perform biped
  locomotion in environments with variable conditions which are not
  previously given to the robot, and plan according to the perception
  of those variable conditions.

\item To globally minimize energy consumption: For a given objective
  in an arbitrary environment, select and perform the gait trajectory
  that minimizes energy consumption.

\end{itemize}


Although non strictly, the order of goals presented has in mind a
growing complexity required to satisfy them. That implies that in some
cases it would be useful to satisfy, at least partially, a previous
one before the one following it.

The last goal proposed, the global minimization of energy consumption,
has not been completely satisfied by any of the works reviewed so far,
and the goals of motion across unstructured environments and planning
of trajectories to attain specific objectives have been taken into
account by few of them. Particularly, in the area of computational
intelligence it is necessary first to address consistently the problem
of planning previous to studying the remaining goals.


Therefore, the specific problem defined is to perform the motion
planning and control of biped locomotion in a computer-simulated
environment using a goal-oriented integrated architecture based on
computational intelligence. The biped model defined is a rigid body
linked model, with the lesser number of joints required to move in a
two-dimensional environment.


The motion planning is understood as the process to detail a motion
task into a sequence of reference actions, while the motion control is
the process that transforms a sequence of reference actions into
control signals applied to robot actuators (applying torque
accordingly to each joint).The computer-simulated environment is
essentially a physics engine in which the dynamical experiments can be
performed, given a model of the system, environmental conditions and a
set of inputs.


% \section{Research Objectives}

% \subsection{General Research Objective}
% To perform the motion planning and control of biped locomotion in a
% computer-simulated environment using a goal-oriented integrated
% architecture based on computational intelligence.

% \subsection{Specific Research Objectives}

% \begin{itemize}
% \item To develop or configure a computer-simulated environment that
%   allows to perform dynamical experiments of motion tasks with the
%   biped
% \item To develop or configure a visualization module that allows to
%   evaluate the performance of the simulations made in the simulation
%   environment
% \item To design and implement algorithmically a goal-based
%   integrated architecture of motion planning and control of biped
%   locomotion and develop the interface with the simulation
%   environment
% \end{itemize}


\section{Main Contributions}
The goal-oriented \sidecomm{control architecture} using neural fields
proposed here, is the main contribution of this work. It shows that
neural fields may be used as a useful component to build motion
controllers, taking into account that they are also biologically
plausible.

In order to be able to move consistently across an unstructured
environment and to be useful for any specific task it is required that
the biped can pursue specific goals. Being able to perform both
planning and control is important for two reasons. The first one is to
allow a uniform and hierarchical implementation across the
architecture, allowing the same adaptation strategies and analysis
tools to be applied to both planning and control systems, and
performing the planning task dynamically in the same time scale in
which the control system is working. The second one is to be able to
change softly from one operation mode to another when goals have
changed, controlling undesired high frequency components of the
reference control signal.

An additional but important contribution, related to the goal of
minimization of \sidecomm{global energy consumption}\!, is the
optimization by means of evolutionary algorithms and neural fields of
the previous work on pseudo-Passive Dynamic Walking controllers (see
\cite{Wisse05How}) for the Simplest Biped Walking model (as proposed
in \cite{Garcia98simplest}). With this contribution, we are advancing
towards what could be the most valuable goal on biped walking control.

The proposed problem faces dynamic biped walking in structured
environments without perturbations. With the contribution on
goal-oriented planning and its integration to motion control using a
computational intelligence approach, and furthermore with the
improvement on energy consumption, it is expected that this work
provides a framework such that the motion across unstructured
environments and the global energy minimization for general
environments will fall in the field of study in short time.

\section{Contents Overview}
The contents of this work are organized as follows:

\begin{itemize}
\item First some preliminaries are presented, including:
  \begin{itemize}
  \item The problem of biped walking and some general approaches to
    solve it;
  \item An introduction to neural fields.
  \item An introduction to evolutionary algorithms.
  \item And finally, a survey of computational intelligence applied to
    biped robotics.
  \end{itemize}
\item Then, a first approach to a neural field control architecture is
  shown. It is used to solve the stability problem for the inverted
  pendulum, and compared against a recurrent neural network
  controller. Also, it is shown a second neural field controller,
  where adaptation using evolutionary algorithms is applied.
\item Later, an extended neural field control and planning
  architecture is developed and applied to the stability problem for
  the Simplest Biped Walking model. In this chapter we show how the
  control architecture may be used at a planning level by changing the
  control policy to be applied, and also how it is able to minimize
  the global energy consumption. The neural field control architecture
  is compared to the linear controller proposed by Wisse et al., and
  to the optimized but not biologically inspired Sliding-mode
  controller proposed by us.
\item Finally, some conclusions are drawn and possibilities for future
  work are presented.
\end{itemize}



